Differential invariant signatures for planar Lie group transformations of images

### 2019-11-06 paper/newpaper.tex
Morning's work

\date{\today}
\title{Differential invariant signatures for planar Lie group transformations of images}
\author{Not-working authors}

@@

\section{Introduction}
\todo[inline]{This is a start. To be completed later.}

@@

Following the treatment in~\cite{OlverCIT}, this leads to Cayley's Omega Process~\cite{Cayley}. We will demonstrate this for the affine group $A(2)$ and its subgroup $SA(2)$.

Consider $\mathbf{x}_i \in \mathbb{R}^2, i=1,2$, so that $\mathbf{x}_i$ can be written in coordinates as $(x_i, y_i)$. Then a pair of such $\mathbf{x}$ lie in $\mathbb{R}^2 \times \mathbb{R}^2$.  Now consider applying an affine transformation to each $\mathbf{x}$ independently: $\mathbf{x}_i \mapsto A \mathbf{x}_i + b$ where $A \in GL(2,\mathbb{R}$ and $b \in \mathbb{R}^2$. Then the Omega Process is a second-order differential operator defined by:

\begin{equation}
\Omega_{ij} = \begin{vmatrix} \frac{\partial}{\partial x_i} & \frac{\partial}{\partial y_i} \\ \frac{\partial}{\partial x_j} & \frac{\partial}{\partial y_j} \end{vmatrix}  = \frac{\partial^2}{\partial x_i \partial y_j} - \frac{\partial^2}{\partial x_j \partial y_i}.
\end{equation}

Under the above affine transformation, $\Omega_{ij} \mapsto \det A^{-1} \Omega_{ij}$. Note that $\det A^{-1}$ has no dependence on the coordinates.

The Omega Process can be applied to pairs of smooth functions as follows:

\begin{equation}
\Omega_{ij} f(x_i, y_i) g(x_j, y_j) = ...
\end{equation}

@@

\noindent where $mbox{tr}$  is the operator that identifies the coordinates of each function $x_i = x_j = x, y_i = y_j = y$.

This can be generalised to $n$ functions and $r-$th order~\cite{Olver} as:

\begin{equation}
\mbox{tr} \left( \Pi_{k=1}^r \Omega_{i_k j_k} \right) f_1(x_1, y_1) f_2(x_2, y_2), \ldots f_n (x_n, y_n)
\end{equation}

Under this formula we end with a factor $\det A^{-r}$ in place of $\det A^{-1}$ that we saw previously. $r$ is referred to as the weight of the partial transvectant. It can also be useful to consider the maximum number of derivatives computed, which we term degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.
For the current case it is sufficient to consider $n$ copies of the same function $f$.

This formula can be applied to generate partial transvectants, as we will now show for the action of two groups: $A(2)$ itself, and its subgroup $SA(2)$.

\todo[inline]{Start here... write out some bits.}

Then computations.

Then the graphs and the weights.

\begin{tabular}{cm{1cm}cm{10cm}}
\centering
Degree & Weight & Diagram & Invariant \\
\hline
2 & 2 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
& $2 f_{xx} f_{yy} - 2 \left(f_{xy}\right)^{2}$ \\

@@

4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (-0.9511,0.3090) {2};
    \node[draw,circle,minimum size=0.2cm] (3) at (-0.5878,-.8090) {3};
    \node[draw,circle,minimum size=0.2cm] (4) at (0.5878,-.8090) {4};
    \node[draw,circle,minimum size=0.2cm] (5) at (0.9511,0.3090) {5};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
    \draw[-] (1) to (4);
    \draw[-] (1) to (5);
\end{tikzpicture}
& $\left(f_{x}\right)^{4} f_{yyyy} - 4 \left(f_{x}\right)^{3} f_{y} f_{xyyy} + 6 \left(f_{x}\right)^{2} \left(f_{y}\right)^{2} f_{xxyy} - 4 f_{x} \left(f_{y}\right)^{3} f_{xxxy} + f_{xxxx} \left(f_{y}\right)^{4}$ \\
\end{tabular}

@@

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SA2.png}
\includegraphics[width=.45\textwidth]{Figs/SA2_signature.png}
\caption{A sample $SA(2)$ Transform and 3D signature.}
\label{fig:SA2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_A2.png}
\includegraphics[width=.45\textwidth]{Figs/A2_signature.png}
\caption{A sample $A(2)$ Transform and 3D signature.}
\label{fig:A2}
\end{figure}

>>>>

\date{\today}
\title{Differential invariant signatures for planar Lie group transformations of images}
\author{Not-working authors }

@@

\section{Introduction}
\todo[inline]{Examples to be numbered like theorems}
\todo[inline]{This is a start. To be completed later.}

@@

We will demonstrate this for the affine group $A(2)$ and its subgroup
$SA(2)$ where, following the treatment in~\cite{OlverCIT}, this leads to
Cayley's Omega Process~\cite{Cayley}.

Consider $\mathbf{x}_i \in \mathbb{R}^2, i=1,2$, so that $\mathbf{x}_i$ can
be written in coordinates as $(x_i, y_i)$. Then a pair of such $\mathbf{x}$ lie in $\mathbb{R}^2 \times \mathbb{R}^2$. Now consider applying the same affine transformation to each $\mathbf{x}$ independently: $\mathbf{x}_i \mapsto A \mathbf{x}_i + b$ where $A \in GL(2,\mathbb{R}$ and $b \in \mathbb{R}^2$. Then the Omega Process is a second-order differential operator defined by: 
\begin{equation}
\Omega_{ij} = \begin{vmatrix} \frac{\partial}{\partial x_i} &
\frac{\partial}{\partial y_i} \\ \frac{\partial}{\partial x_j} &
\frac{\partial}{\partial y_j} \end{vmatrix} = 
\frac{\partial^2}{\partial x_i
\partial y_j} - \frac{\partial^2}{\partial x_j \partial y_i}.
\end{equation}

Under the above affine transformation, $\Omega_{ij} \mapsto \det A^{-1} \Omega_{ij}$. Note that $\det A^{-1}$ has no dependence on the coordinates. 
The Omega Process can be applied to pairs of smooth functions as follows, writing $\frac{\partial f}{\partial x}$ as $f_x$:

\begin{equation}
\Omega_{ij} f(x_i, y_i) g(x_j, y_j) = f_{x_i}g_{y_j} - f_{y_i}g_{x_j}
\end{equation}

@@

\noindent where $\mbox{tr}$  is the operator that identifies the coordinates of each function $x_i = x_j = x, y_i = y_j = y$.

This can be generalised to $n$ functions $f^{(1)}(x_1, y_1), \ldots, f^{(n)}(x_n, y_n)$ and $r$-th order~\cite{Olver} to define a partial transvectant as:

\begin{equation}
\mbox{tr} \left[\left( \Pi_{k=1}^r \Omega_{i_k j_k} \right) f^{(1)}(x_1, y_1)
f^{(2)}(x_2, y_2), \ldots f^{(n)} (x_n, y_n)\right]
\end{equation}
where $i_k \neq j_k \in \{1, \ldots, n\}$.

Under this formula we end with a factor $\det A^{-r}$ in place of $\det
A^{-1}$ that we saw previously, when all $n$ copies of $\mathbb{R}^2$ are
transformed by the same affine function $\mathbf{x} \mapsto A\mathbf{x} +
\mathbf{b}$. $r$ is referred to as the weight of the partial
transvectant. It can also be useful to consider the maximum order of
derivative in the transvectant, which we term degree, e.g., $f_{xy}^2 f_x
f_y$ has degree 2. Note also that the individual pairwise Omega processes commute.

For the current case it is sufficient to consider $n$ copies of the same
function $f$. If we represent each copy (of $\mathbb{R}^2$) as a node on a
undirected graph, and each Omega process $\Omega_{ij}$ as an edge joining
nodes $i$ and $j$, a partial transvectant can be compactly represented in a
graphical form.

For example, consider a weight $2$ transvectant defined as
\begin{align*}
    \mbox{tr} &\left[(\Omega_{12}\Omega_{12}) f(x_1, y_1)f(x_2, y_2)\right] \\
    &= \mbox{tr}\left[(\partial_{x_1 y_2} - \partial_{x_2y_1})(\partial_{x_1 y_2} -
    \partial_{x_2y_1})f(x_1, y_1)f(x_2, y_2)\right]\\
    &= \mbox{tr}\left[\left(\partial_{x_1 y_2} - \partial_{x_2 y_1}\right)\left(f_x(x_1, y_1)f_y(x_2, y_2) - f_y(x_1, y_1)f_x(x_2, y_2)\right)\right] \\
    &= \mbox{tr}\left[f_{xx}(x_1, y_1)f_{yy}(x_2, y_2) - f_{xy}(x_1, y_1)f_{xy}(x_2, y_2)\right] \\
    &= f_{xx}(x, y) f_{yy}(x, y) - f_{xy}(x, y)^2
\end{align*}
This is a degree $2$ partial transvectant as it involves derivatives of $f$
up to second order, and can be represented graphically by the following
diagram,
\begin{center}
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
\end{center}
where the double edge between nodes $1$ and $2$ corresponds to the two copies of $\Omega_{12}$.
Table~\ref{tab:transvectants} shows all possible nonzero partial
transvectants of $A_2$ up to weight $4$ that can be generated in this way.
Note that the weight is the number of edges in the graph.
\begin{table}
\centering
\begin{tabular}{cccp{11cm}}
Degree & Weight & Diagram & Partial transvectant\\
\hline
2 & 2 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
& $2 f_{xx} f_{yy} - 2 \left(f_{xy}\right)^{2}$ \\

@@

4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.2cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.2cm] (2) at (-0.9511,0.3090) {2};
    \node[draw,circle,minimum size=0.2cm] (3) at (-0.5878,-.8090) {3};
    \node[draw,circle,minimum size=0.2cm] (4) at (0.5878,-.8090) {4};
    \node[draw,circle,minimum size=0.2cm] (5) at (0.9511,0.3090) {5};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
    \draw[-] (1) to (4);
    \draw[-] (1) to (5);
\end{tikzpicture}
& $\left(f_{x}\right)^{4} f_{yyyy} - 4 \left(f_{x}\right)^{3} f_{y} f_{xyyy} + 6 \left(f_{x}\right)^{2} \left(f_{y}\right)^{2} f_{xxyy} - 4 f_{x} \left(f_{y}\right)^{3} f_{xxxy} + f_{xxxx} \left(f_{y}\right)^{4}$ \\
\end{tabular}
\caption{Details and graphical representations of all partial transvectants up to weight $4$ of copies of a function $f$ under a transformation in $A(2)$.}
\label{tab:transvectants}
\end{table}

\subsection{Signatures of $SA(2)$ and $A(2)$}
Recall that invariants of a supergroup are automatically invariants of any of
its subgroups. For example, invariants of $A(2)$ are automatically invariants of $SA(2)$, but not necessarily vice versa. Any of the partial transvectants found in Table~\ref{tab:transvectants} are invariants of $SA(2)$ (as the determinant is $1$) but only up to scale factor ($\det A^{-r}$) of $A(2)$.

Hence, we can form a signature for $SA(2)$ by choosing any combination of
partial transvectants. We'll choose a signature with three components so that
it can be visualised as a three-dimensional surface; note that this will not
be complete and may well have a \emph{very large bad set}. For example, we can choose the function value together with the first two partial transvectants to form the signature
\begin{equation*}
    \left(f, f_{xx}f_{yy}-f_{xy}^2, f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2\right)
\end{equation*}

\subsubsection{Example}
Consider as our image, the function $f(x, y) = 0.5 - 0.2x - 0.3y - 0.05x^2 +
0.03xy + 0.04y^2 + 0.03y^3 +0.001x^3 - 0.0015x^2y + 0.002xy^2 - 0.0005y^3$
defined on $[0, 1]\times[0, 1]$.
\todo[inline]{remake figure with modified matrix. Also modify function so it has some features like a critical point}
We'll subject our image to a special affine transformation:
\begin{equation*}
\mathbf{x} \mapsto \frac{1}{1.1}\begin{bmatrix} 0.6875 & 1.1 \\ -0.6 & 0.8 \end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
\end{equation*}
Fig.~\ref{fig:SA2} shows the function as a greyscale image, the transformed
function, and the signatures of both the untransformed and transformed
images.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SA2.png}
\includegraphics[width=.45\textwidth]{Figs/SA2_signature.png}
\caption{\todo[inline]{add extra crap} A sample $SA(2)$ Transform and 3D signature. Mention what's going on with the grid.}
\label{fig:SA2}
\end{figure}

\subsubsection{Moving on to $A(2)$, \ldots}
Using the same function as an example, 
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_A2.png}
\includegraphics[width=.45\textwidth]{Figs/A2_signature.png}
\caption{A sample $A(2)$ Transform and 3D signature.}
\label{fig:A2}
\end{figure}

\todo[inline]{visual obstruction}

@@
